Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-27 TACAS2018/paper.tex
changes synthesis section

The CEGIS-OR loop for this approach is described next.
We make use of two abstractions corresponding to two
specifications: $\phi_{synth}$ used by the synthesiser and
$\phi_{static}$ which is given to the optimiser. 
Both specifications are updated during execution and new abstractions are constructed for each.
Let us define the properties 
\begin{align*}
\phi_{\langle C \rangle}&:= \mat{K}=\mat{K}_{\langle C \rangle} \wedge \mat{L}=\mat{L}_{\langle C \rangle}\\
\phi_{dynamics}^{T_d}&:= \{\mat{A}_d, \mat{B}_d, \mat{C}_d \mid \mat{A}_d=e^{\mat{A}T_d}\}
\end{align*}
$\phi_{\langle C \rangle}$ defines the word length of the digital controller ($\mathbb{R}_{\langle C \rangle}$)
and $\phi_{dynamics}^{T_d}$ defines the sampling time ($T_d$) and its effect on the system dynamics (as per Section \ref{sec:continuous_time_accel}).
We start our loop with 
\begin{align*}
\phi_{synth}&:=\phi_{dynamics}^{T_d} \wedge \phi_{stability} \wedge \phi_{input} \wedge \phi_{init}\\
\phi_{static}&:=\phi_{synth} \wedge \phi_{\langle C \rangle}.
\end{align*}
These specifications will vary as the loop progresses.  However, at all times $\phi_{static} \models \phi_{synth}$.
Let us see how the loop updates these specifications.

\paragraph{1. {\sc abstract}} creates an abstraction from the initial specification as follows:
\begin{itemize}
\item Select a sample time ($T_d$) that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep2}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. 
%The transformation matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:cof_fwl_stability}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
When a counterexample is found, we add it to $\phi_{synth} = \phi_{synth} \wedge \phi_{safety}^k$, 
where k is the iteration found for the counterexample.  The ${\sc Abstract}$
phase will then construct a abstraction (in the form of a program) that meets
this property (unsoundly since the program uses FWL).

%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision ($\phi_{\langle P \rangle}$), or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

%\item
\paragraph{4. {\sc Optimise}} refines the candidate solution towards
a local optimum.  We seek to minimize the distance of the farthest point in the
reach space to the closest point to it in the safe region
\begin{equation}
e=max(\{min(\{\vec{x}-\vec{g} \mid \vec{g}\in G\}) \mid \vec{x} \in \hat{X}_\infty \}.
\label{eq:objective}
\end{equation}

Our optimization is iterative, following two stages.  The first stage seeks to
make the dynamics as convergent as possible, since convergent dynamics are
less likely to stray outside of the safe region (we say less likely because the
change in the eigenvectors changes the shape of the space and may cause some 
trajectories to go further).

In this case, we take the largest eigenvalue and reduce it by a factor $f$
that still satisfies the controller constraints ($f \overline{\lambda} \models \phi_{static}$).
The resulting observer controller is then verified against the safety specification $\phi_{safety}$
and the objective $e$ is measured (Equation \eqref{eq:objective}).
If for each iteration $k$, the objective $e_k<e_{k-1}$, we continue to iterate until 
$e_k \leq 0$ (i.e. we found a safe controller), otherwise we move to the second stage.

In the second stage, we take the complex eigenvalue
with the largest angle ($max\{|tan^{-1}\left(\frac{im(\lambda)}{re(\lambda)}\right)| \mid \lambda \in \Lambda\}$)
and reduce its imaginary component only  ($re(\lambda) + f im(\lambda) \models \phi_{static}$).
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and has a
smaller rotation, which facilitates meeting the safety specification $\phi_{safety}$
(we note that real convergent eigenvalues are trivially safe).

We denote these sets of rules as $\phi_{objective}$ (which varies depending on which optimization stage we're in).
At each optimizing iteration, the modeller updates $\phi_{static} = \phi_{static} \wedge \phi_{objective}$% \wedge \phi_{safety}^{k'} $ where $k' \geq k$.

If we cannot optimize the candidate, we signal the {\sc Abstract} phase to
refine $\phi_{synth}$ and request {\sc Synthesise} to produce a new
candidate.  Otherwise the optimised candidate solution is given back to
{\sc Verify Synthesis}.

@@

    ([xshift=5em]model1.south) edge node[align=center,xshift=-.65cm] {$\langle P \rangle$} ([xshift=5em]synth.north) node[align=left,yshift=-0.7cm,xshift=.8cm] {{$\phi_{stability}$}\\{$\phi_{safety}$}\\ {$\phi_{init}^K$}}

>>>>

%The CEGIS-OR loop for this approach is described next.
Throughout the CEGIS-OR loop we make use of a specification $\phi$
which encompasses the abstraction of the system's dynamics.
Note that, given that the system's abstraction gets refined
in each iteration, $\phi$ will also be adjusted appropriately.
Initially, $\phi$ is given by: 
$\phi=\phi_{stability} \wedge \phi_{input} \wedge \phi_{init}$.

%% We make use of two abstractions corresponding to two
%% specifications: $\phi_{synth}$ used by the synthesiser and
%% $\phi_{static}$ which is given to the optimiser. 
%% While both specifications are updated during execution, for now we describe their
%% meaning at the beginning of the synthesis process.
%% If we denote by $\phi_{dynamics}^{T_d}$ the discretized system's dynamics with respect to the
%% sampling time ($T_d$),
%% $\phi_{dynamics}^{T_d}:= \{\mat{A}_d, \mat{B}_d, \mat{C}_d \mid \mat{A}_d=e^{\mat{A}T_d}\},$
%% then initially $\phi_{synth}$ and $\phi_{static}$ are equal to:

Note that there is a major difference between
the way $\phi$ is being used in different phases in Figure~\ref{fig:CEGIS-OR}: while the
{\sc synthesise} phase checks $\phi$ by using the plant's precision
$\mathbb{R}_{\langle P \rangle}$ for the controller, observer and
plant, the {\sc model} and {\sc verify} phases use reals to represent
the plant and $\mathbb{R}_{\langle C \rangle}$ for the controller and
the observer. The reason for this is that, as we want the
synthesis to be fast, we allow it to generate unsound results
and we restore soundness during the verification and modeling stages.
%% While these specifications are updated throughout the synthesis process (as appropriately described below), at all times $\phi_{static} \models \phi_{synth}$.
%Let us see how the loop updates these specifications.

Next, we describe the CEGIS-OR loop in Figure~\ref{fig:CEGIS-OR}:
\paragraph{1. {\sc abstract}} computes an initial version of $\phi$
and decides the sample time $T_d$,
the plant's precision $\mathbb{R}_{\langle P \rangle}$ and the controller's precision
$\mathbb{R}_{\langle C \rangle}$.
 %width for the bit vectors
 %for representing the controller and the plant, respectively.
%
%For this creates an abstraction from the initial specification as follows:
\begin{itemize}
\item Select a sample time ($T_d$) that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\note{implementation and abstraction?}
\item Select initial precisions $\mathbb{R}_{\langle P \rangle}$ and $\mathbb{R}_{\langle C \rangle}$
  %width for the bit vectors
  for representing the controller and the plant, respectively. These will be increased as needed.
 % $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed. 
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$. % using the truncation described in Section~\ref{sec:numeric_rep2}.
\item Transform the system into canonical form and add nondeterminism for the errors
as described in Section~\ref{sec:cof_fwl_stability}. 
%The transformation matrices will be needed to obtain an FWL compliant controller and observer.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:cof_fwl_stability}.
\item Compute necessary characteristic polynomials. 
%% \item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
%% as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
%% \item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
%% as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification $\phi_{stability}$ based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time %%(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\note{could not find perturbation error anywhere}
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
%% \item Under certain circumstances (i.e. when {\sc optimize} did not manage to improve the s) When a counterexample is found, we add it to $\phi_{synth} = \phi_{synth} \wedge \phi_{safety}^k$, 
%% where k is the iteration found for the counterexample.  The ${\sc Abstract}$
%% phase will then construct a abstraction (in the form of a program) that meets
%% this property (unsoundly since the program uses FWL).
\end{itemize}
\note{$\phi_{safety}^k$ is undefined!!!}
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
If there is no candidate solution it signals {\sc Abstract}, 
which may choose to increase the plant's precision or return a synthesis failure.
If {\sc synthesize} generates a candidate solution, it is fed to the {\sc verify} chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies that the candidate controller
and observer obey $\phi$ (remember that we now use
$\mathbb{R}_{\langle C \rangle}$ for the controller's representation
and reals for the plant).  If the verification fails, the candidate is
fed to the {\sc Optimise} phase.  If the candidate is valid, we feed
it to the next {\sc verify} stage, {\sc Discrete Time}.

%\item
\paragraph{4. {\sc Optimise}} optimises the failed candidate solution.
We seek to minimize the distance of the farthest point in the
reach space to the closest point to it in the safe region
\begin{equation}
e=max(\{min(\{\vec{x}-\vec{g} \mid \vec{g}\in G\}) \mid \vec{x} \in \hat{X}_\infty \}).
\label{eq:objective}
\end{equation}

Our optimization is iterative, following two stages.  The first stage
seeks to make the dynamics as convergent as possible, since convergent
dynamics are less likely to stray outside of the safe region (we say
less likely because the change in the eigenvectors changes the shape
of the space and may cause some trajectories to go further). For this
purpose, we take the largest eigenvalue and reduce it by a factor $f$
that still satisfies the controller constraints ($f \overline{\lambda}
\models \phi$).  The resulting observer and controller are then
verified against the safety specification $\phi_{safety}$ and the
objective $e$ is measured (Equation \eqref{eq:objective}).  If for
each iteration $k$, the objective $e_k<e_{k-1}$, we continue to
iterate until $e_k \leq 0$ (i.e. we found a safe controller),
otherwise we move to the second stage.

In the second stage, we take the complex eigenvalue
with the largest angle ($max\{|tan^{-1}\left(\frac{im(\lambda)}{re(\lambda)}\right)| \mid \lambda \in \Lambda\}$)
and reduce its imaginary component only  ($re(\lambda) + f im(\lambda) \models \phi$).
The process ends when no imaginary part can be further reduced.
The resulting closed loop form converges faster than the original and has a
smaller rotation, which facilitates meeting the safety specification $\phi_{safety}$.
%(we note that real convergent eigenvalues are trivially safe).

%% We denote these sets of rules as $\phi_{objective}$ (which varies depending on which optimization stage we're in).
%% At each optimizing iteration, the modeller updates $\phi_{static} = \phi_{static} \wedge \phi_{objective}$% \wedge \phi_{safety}^{k'} $ where $k' \geq k$.

If we cannot optimize the candidate, we signal the {\sc Abstract} phase to
refine $\phi$ and request {\sc Synthesise} to produce a new
candidate.  Otherwise the optimised candidate solution is given back to
{\sc Verify Synthesis}.

@@

   ([xshift=5em]model1.south) edge node[align=center,xshift=-.7cm] {$\mathbb{R}_{\langle P \rangle}$} 
([xshift=5em]synth.north)
   node[align=left,yshift=-0.7cm,xshift=.5cm] {{$\phi$}\\{$\phi_{init}^K$}} %{$\phi_{safety}$}\\ 
